---
title: "Homework 6"
author: "Julia Thompson"
date: "11/20/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

We load and clean the data for regression analysis, and find that there are no missing values in any of the columns. We also recode and convert _babysex_, _frace_, _malform_, and _mrace_ to factor variables. 

```{r}
birthweight = read_csv("./data/birthweight.csv") 

# check for missing data

ismissing = map(.x = birthweight, ~ sum(is.na(.x))) %>%  # why do you need the ~ ?!?
  bind_cols()

# male = 1, female = 2
# 1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown
# 0 = absent, 1 = present
# 1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other

# Recode variables to make sense

birthweight = birthweight %>% 
  mutate(
    babysex = recode(babysex, `1` = "male", `2` = "female"),
    frace = recode(frace, `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other", `9` = "unknown"),
    malform = recode(malform, `0` = "absent", `1` = "present"),
    mrace = recode(mrace,  `1` = "white", `2` = "black", `3` = "asian", `4` = "puerto rican", `8` = "other")
  )

# change the above recoded variables into factors with levels corresponding to the above order

birthweight = birthweight %>% 
  mutate(
    babysex = factor(babysex, levels = c("male", "female")),
    frace = factor(frace, levels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
    malform = factor(malform, levels = c("absent", "present")),
    mrace = factor(mrace, levels = c("white", "black", "asian", "puerto rican", "other"))
  )
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

First, we used a hypothesized structure consulting literature for the factors that underly birthweight to reduce our number of potential predictors. We will continue with the following: gender (_babysex_), mother's BMI (_ppbmi_), mother's height (_mheigth_), cigarettes smoked per day (_smoken_), gestational age (_gaweeks_), mother's age (_momage_), and mother's race (_mrace_). 

We will then choose a final model using stepwise backwards elimination. We begin with a model including all predictors chosen above, and then one by one eliminate non-significant predictors. We use a p-value of 0.15 here, because 

```{r}
# create a model for birthweight
# first, get testing and training datasets

# cv_df =
#   crossv_mc(birthweight, 100) %>% 
#   mutate(
#     train = map(train, as_tibble),
#     test = map(test, as_tibble))

# hypothesize that mother's race might have an interaction with weight


```



